---
title: "GE culture paper"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---


```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```



# Comments
* A Rmd file must be end as .Rmd and not rmd
* Do not use data.table and dplyr, just use one to unify the syntax.  I recommend dplyr by far
* Do not use plyr which is an old library and superseded by r
* R chunks must begin by ```{r} to be recognized and not ```r
* Use readr library
* No need to export from Excel to csv.  Use the readxl package (if you read from csv or tsv, use readr)
* Do not use factors as much as possible, this is not really useful

# Figures for the paper
## Load libraries

```{r}
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(readxl)
library(stringr)
```


## Import csv

```{r}
# culturesGE <- fread("culturesGE.csv", stringsAsFactors = TRUE)

culturesGE  <- readxl::read_excel(path=file_main, sheet="Data S1")
```
## Order categorical data in "class"

```{r}
culturesGE$class <- factor(culturesGE$class, levels = ordered(c("Bacillariophyceae", "Bacillariophyta_X", "Coscinodiscophyceae", "Mediophyceae", "Chlorophyceae", "Mamiellophyceae", "Prasinophyceae", "Cryptophyceae", "Coccolithophyceae", "Prymnesiophyceae", "Dictyochophyceae", "Dinophyceae", "Chrysophyceae", "Pelagophyceae")))
```

## Separate by diatom and non-diatom

```{r}
culturesGE_diatoms <- culturesGE %>% dplyr::filter(group=="diatom")

culturesGE_non_diatoms <- culturesGE %>%  dplyr::filter(group=="non diatom")
```

## Create palette

```{r}
GE_class_pallette <- c(Bacillariophyceae = "#660066", 
                       Bacillariophyta_X = "#ff748c", 
                       Coscinodiscophyceae = "#08020d", 
                       Mediophyceae = "#d896ff", 
                       Chlorophyceae = "#3cb44b", 
                       Mamiellophyceae = "#004c00", 
                       Prasinophyceae = "#74c476", 
                       Cryptophyceae = "#F0E442", 
                       Coccolithophyceae = "#00004c", 
                       Prymnesiophyceae = "#2b8cbe", 
                       Dictyochophyceae = "#de2d26", 
                       Dinophyceae = "#800000", 
                       Chrysophyceae = "#fcae91", 
                       Pelagophyceae = "#fb6a4a")
```

## Plot bar graphs Isolation

```{r}
theme_set(theme_bw())

isolation_method_plot_diatoms <- 
  ggplot(culturesGE_diatoms, aes(isolation_method)) +
  geom_bar(aes(fill=class), width = 0.5) + 
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1, size=12)) + 
  facet_wrap(~ sampling_substrate) + 
  xlab("Isolation method") +
  ylab("Number of strains") +
  scale_fill_manual(values = GE_class_pallette)

isolation_method_plot_diatoms
  
isolation_method_plot_non_diatoms <- 
  ggplot(culturesGE_non_diatoms, aes(isolation_method)) + 
  geom_bar(aes(fill=class), width = 0.5) + 
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1, size=12)) + 
  facet_wrap(~ sampling_substrate) + 
  xlab("Isolation method") +
  ylab("Number of strains") +
  scale_fill_manual(values = GE_class_pallette)

isolation_method_plot_non_diatoms
```

## Filter by Amundsen samples

```{r}
culturesGE_AMUNDSEN <- culturesGE %>%
  filter(sampling_cruise=="Green Edge Amundsen")
```

## Plot bar graphs Amundsen


```{r fig.height=10, fig.width=10}
isolation_method_AM_plot_station <- ggplot(culturesGE_AMUNDSEN, aes(isolation_method)) + 
  geom_bar(aes(fill=class), width = 0.7) + 
  theme(axis.text.x = element_text(angle=65, vjust=1, hjust=1, size=10)) + 
  facet_wrap(~ sampling_station) + 
  xlab("Isolation method") +
  ylab("Number of strains") +
  scale_fill_manual(values = GE_class_pallette)

isolation_method_AM_plot_station 

isolation_method_AM_plot_depth <-  ggplot(culturesGE_AMUNDSEN, aes(isolation_method))+
  geom_bar(aes(fill=class), width = 0.7) + 
  theme(axis.text.x = element_text(angle=65, vjust=1, hjust=1, size=10)) + 
  facet_wrap(~ sampling_depth_values) + 
  xlab("Isolation method") +
  ylab("Number of strains") +
  scale_fill_manual(values = GE_class_pallette)

isolation_method_AM_plot_depth
```

# Tables in Latex format

* Use the xtable package

Note : italics are done by encosing between {}

## Load libraries

```{r}

library(xtable)
```

## Define function to format the tables
```{r, eval=FALSE}
sanitize.italics <- function(str) {
  str_replace_all(str, c("_" = "\\\\_", 
                         "\\{" = "\\\\textit{", 
                         "°" = "\\\\degree",
                         "X" = "\\\\cellcolor{gray}"))
}


path_table <- function(file_table) {str_c("../overleaf_paper_ge_cultures_2.0/tables/", file_table)}
file_main = "GE_cultures_Tables.xlsx"
```

# Table of genera (Table 1)
```{r, eval=FALSE}
table <- readxl::read_excel(path=file_main, sheet="Table 1")
table <- xtable::xtable(table, label="table:genera", 
                        caption='Number of strains obtained from water and ice samples for each genus', 
                        align = c("l", "l", "l", "c", "c"),
                        digits=0)
print(table,  scalebox = 0.75, 
      caption.placement = "top",
      include.rownames = FALSE, 
      file=path_table("table_genera.tex"),
      sanitize.text.function = sanitize.italics)
```

# Table of authorithies (Table S1)
```{r, eval=FALSE}
table <- readxl::read_excel(path=file_main, sheet="Table S1")
table <- xtable::xtable(table, label="table:authorities", 
                        caption='List of taxonomy authorities', 
                        align = c("l", "l", "l", "l"),
                        digits=0)
print(table,  scalebox = 0.75, 
      caption.placement = "top",
      include.rownames = FALSE, 
      file=path_table("table_authorities.tex"),
      sanitize.text.function = sanitize.italics)
```
