---
title: "GE culture paper"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---


```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```



# Comments for Catherine
* A Rmd file must be end as .Rmd and not rmd
* Do not use data.table and dplyr, just use one to unify the syntax.  I recommend dplyr by far
* Do not use plyr which is an old library and superseded by r
* R chunks must begin by ```{r} to be recognized and not ```r
* Use readr library
* No need to export from Excel to csv.  Use the readxl package (if you read from csv or tsv, use readr)
* Do not use factors as much as possible, this is not really useful

# Load libraries

```{r}
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(readxl)
library(stringr)
library(dvutils)
library(pr2database)
```

# Define variables
```{r}
file_main = "GE_cultures_Tables.xlsx"
```



# BLAST analysis

## Process BLAST file

```{r, eval=FALSE}
blast_18S_reformat("GE_cultures_18S.blast_pruned.tsv")
```

## Figures BLAST for the paper

### Import csv

```{r}
# culturesGE <- fread("culturesGE.csv", stringsAsFactors = TRUE)

culturesGE_Blast  <- readxl::read_excel(path=file_main, sheet="Data S2")
```

### BLAST figure

```{r}
GE_class_pallette <- c(Bacillariophyceae = "#660066", 
                       Bacillariophyta = "#ff748c", 
                       Coscinodiscophyceae = "#08020d", 
                       Mediophyceae = "#d896ff", 
                       Chlorophyceae = "#3cb44b", 
                       Mamiellophyceae = "#004c00", 
                       Pyramimonadales = "#74c476", 
                       Cryptophyceae = "#F0E442", 
                       Coccolithophyceae = "#00004c", 
                       Prymnesiophyceae = "#2b8cbe", 
                       Dictyochophyceae = "#de2d26", 
                       Dinophyceae = "#800000", 
                       Chrysophyceae = "#fcae91", 
                       Pelagophyceae = "#fb6a4a")

  culturesGE_Blast_table <- culturesGE_Blast %>% 
    mutate (status = case_when (hit_top_pct_identity < 100 ~ "undetected",
                                hit_cult_pct_identity < 100 ~ "detected - uncultured",
                                TRUE ~ "cultured")) %>% 
    group_by(status) %>% 
    tally()
  
    knitr::kable(culturesGE_Blast_table)
 
  ggplot(culturesGE_Blast) + 
  geom_histogram(aes(x=hit_top_pct_identity), fill= "black", alpha=0.8) +
  # geom_histogram(aes(x=hit_cult_pct_identity), fill= "blue", alpha=0.3) +
  xlab("% Identity to top hit")  +
  scale_x_reverse() 
  # ggtitle("Red is top hit, blue top hit to culture")
  
  culturesGE_Blast <- culturesGE_Blast %>% 
    mutate(cluster_label = str_c(hit_pr2_division, hit_pr2_class,  sep=" - "))
  
  ggplot(culturesGE_Blast, aes(y=hit_cult_pct_identity, 
                               x=cluster_label, 
                               color=hit_pr2_class)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(size=2, height=0, width=0.2) +
  theme_bw() +
  # stat_summary(fun.y = "mean", colour = "black", size = 2, geom = "point", shape=22) +
  scale_color_manual(values = GE_class_pallette) +
  theme(axis.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)) +
  ylab("% Identity to best cultured hit") + xlab("") +
  scale_y_reverse() +
  scale_x_discrete(limits = rev(levels(as.factor(culturesGE_Blast$cluster_label)))) +
  coord_flip() +
  theme(legend.position = "none") 
  

 ggplot(culturesGE_Blast, aes(x=hit_top_pct_identity, y=hit_cult_pct_identity, color=hit_pr2_class)) + 
  geom_jitter(size=3, width=0.01, height=0.01)+
  scale_y_reverse()  +
  scale_x_reverse() +
  xlab("% Identity to top hit") + 
  ylab("% Identity to top culture hit")  +
  scale_color_manual(values = GE_class_pallette) 
```


# Figures Catherine for the paper

## Import csv

```{r}
# culturesGE <- fread("culturesGE.csv", stringsAsFactors = TRUE)

culturesGE  <- readxl::read_excel(path=file_main, sheet="Data S1") %>% 
  mutate(sampling_depth = as.numeric(stringr::str_replace(sampling_depth_values,"m",""))) 
```
## Order categorical data in "class"

```{r}
culturesGE$class <- factor(culturesGE$class, levels = ordered(c("Bacillariophyceae", "Bacillariophyta_X", "Coscinodiscophyceae", "Mediophyceae", "Chlorophyceae", "Mamiellophyceae", "Pyramimonadales", "Cryptophyceae", "Coccolithophyceae", "Prymnesiophyceae", "Dictyochophyceae", "Dinophyceae", "Chrysophyceae", "Pelagophyceae")))
```

## Separate by diatom and non-diatom

```{r}
culturesGE_diatoms <- culturesGE %>% dplyr::filter(group=="diatom")

culturesGE_non_diatoms <- culturesGE %>%  dplyr::filter(group=="non diatom")
```

## Create palette

```{r}
GE_class_pallette <- c(Bacillariophyceae = "#660066", 
                       Bacillariophyta_X = "#ff748c", 
                       Coscinodiscophyceae = "#08020d", 
                       Mediophyceae = "#d896ff", 
                       Chlorophyceae = "#3cb44b", 
                       Mamiellophyceae = "#004c00", 
                       Pyramimonadales = "#74c476", 
                       Cryptophyceae = "#F0E442", 
                       Coccolithophyceae = "#00004c", 
                       Prymnesiophyceae = "#2b8cbe", 
                       Dictyochophyceae = "#de2d26", 
                       Dinophyceae = "#800000", 
                       Chrysophyceae = "#fcae91", 
                       Pelagophyceae = "#fb6a4a")
```

## Plot bar graphs Isolation

```{r}
theme_set(theme_bw())

isolation_method_plot_diatoms <- 
  ggplot(culturesGE_diatoms, aes(isolation_method)) +
  geom_bar(aes(fill=class), width = 0.5) + 
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1, size=12)) + 
  facet_wrap(~ sampling_substrate) + 
  xlab("Isolation method") +
  ylab("Number of strains") +
  scale_fill_manual(values = GE_class_pallette)

isolation_method_plot_diatoms
  
isolation_method_plot_non_diatoms <- 
  ggplot(culturesGE_non_diatoms, aes(isolation_method)) + 
  geom_bar(aes(fill=class), width = 0.5) + 
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1, size=12)) + 
  facet_wrap(~ sampling_substrate) + 
  xlab("Isolation method") +
  ylab("Number of strains") +
  scale_fill_manual(values = GE_class_pallette)

isolation_method_plot_non_diatoms
```

## Filter by Amundsen samples

```{r}
culturesGE_AMUNDSEN <- culturesGE %>%
  filter(sampling_cruise=="Green Edge Amundsen") %>% 
  tidyr::replace_na(list(sampling_depth= 0))
```

## Plot bar graphs Amundsen


```{r fig.height=6, fig.width=10}
isolation_method_AM_plot_station <- ggplot(culturesGE_AMUNDSEN, aes(isolation_method)) + 
  geom_bar(aes(fill=class), width = 0.7) + 
  theme(axis.text.x = element_text(angle=65, vjust=1, hjust=1, size=10)) + 
  facet_wrap(~ sampling_station) + 
  xlab("Isolation method") +
  ylab("Number of strains") +
  scale_fill_manual(values = GE_class_pallette)

isolation_method_AM_plot_station 

isolation_method_AM_plot_depth <-  ggplot(culturesGE_AMUNDSEN, aes(isolation_method))+
  geom_bar(aes(fill=class), width = 0.7) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle=65, vjust=1, hjust=1, size=10)) + 
  facet_wrap(~ cut(sampling_depth, c(0,10,20,50), right=FALSE)) + 
  xlab("Isolation method") +
  ylab("Number of strains") +
  scale_fill_manual(values = GE_class_pallette)

isolation_method_AM_plot_depth
```

# Tables in Latex format

* Use the xtable package

Note : italics are done by encosing between {}

## Load libraries

```{r}

library(xtable)
```

## Define function to format the tables
```{r, eval=FALSE}
sanitize.italics <- function(str) {
  str_replace_all(str, c("_" = "\\\\_", 
                         "\\{" = "\\\\textit{", 
                         "°" = "\\\\degree",
                         "X" = "\\\\cellcolor{gray}"))
}


path_table <- function(file_table) {str_c("../paper_ge_cultures_overleaf_2.0/tables/", file_table)}

```

# Table of genera (Table 1)
```{r, eval=FALSE}
table <- readxl::read_excel(path=file_main, sheet="Table 1")
table <- xtable::xtable(table, label="table:genera", 
                        caption='Number of strains obtained from water and ice samples for each genus', 
                        align = c("l", "l", "l", "c", "c"),
                        digits=0)
print(table,  scalebox = 1, 
      caption.placement = "top",
      include.rownames = FALSE, 
      file=path_table("table_genera.tex"),
      sanitize.text.function = sanitize.italics)
```

# Table of authorithies (Table S1)
```{r, eval=FALSE}
table <- readxl::read_excel(path=file_main, sheet="Table S1")
table <- xtable::xtable(table, label="table:authorities", 
                        caption='List of taxonomy authorities', 
                        align = c("l", "l", "l", "l"),
                        digits=0)
print(table,  scalebox = 0.75, 
      caption.placement = "top",
      include.rownames = FALSE, 
      file=path_table("table_authorities.tex"),
      sanitize.text.function = sanitize.italics)
```
